---
title: "TCR simulations - sample size calculations"
output: html_notebook
---

I've run the TCR simulations with several different starting points (random initialisation). I've then randomly sampled different numbers of simulated TCRs from these to 
compare sample sizes.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(viridis)
library(reshape2)
library(lettercase)
source("~/Dropbox/R_sessions/GGMike/theme_mike.R")
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
col.names <- c("TRA_V",	"TRA_J",	"TRA_len", "TRA_nt", "TRA_entropy", "TRA.Misc", "TRA_AA",	"TRA_Peptide",	"TRA_Valid",	"TRA_CDR1_entropy",	"TRA_CDR1",
               "TRA_CDR2_entropy",	"TRA_CDR2",
               "TRA_CDR3_entropy",	"TRA_CDR3",	"TRB_V",	"TRB_J",	"TRB_len", "TRB_nt",	"TRB_entropy",	"TRB_AA",	"TRB_Peptide",	"TRB_Valid",	"TRB_CDR1_entropy",	"TRB_CDR1",
               "TRB_CDR2_entropy",	"TRB_CDR2",	"TRB_CDR3_entropy",	"TRB_CDR3",	"Valid.TCR")
tcr.dir <- "~/CI_filesystem/mnt/scratchb/jmlab/morgan02/TCR_simulations/results/"
sample.size.files <- list.files(tcr.dir, pattern="TCRsim")
sd.list <- list()

for(x in seq_along(sample.size.files)){
  s.size <- unlist(strsplit(sample.size.files[x], split="-", fixed=TRUE))[1]
  s.df <- read.table(paste0(tcr.dir, sample.size.files[x]), sep="\t", header=FALSE, stringsAsFactors=FALSE)
  colnames(s.df) <- col.names
  s.df$Sample.Size <- as.numeric(gsub(s.size, pattern="Sample_", replacement=""))
  sd.list[[s.size]] <- s.df
}

tcr.sims <- do.call(rbind.data.frame,
                    sd.list)
tcr.sims$Sample.Size <- factor(tcr.sims$Sample.Size,
                               levels=c(10, 100, 500, 1000, 5000, 10000, 20000, 50000, 100000))
```

First, what % of TCRs are valid?

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(tcr.sims, aes(x=Sample.Size, fill=Valid.TCR)) +
  geom_bar(position='fill') +
  theme_mike() +
  scale_fill_Publication() +
  labs(x="Sample Size", y="Proportion valid TCRs")

ggsave(filename="~/Dropbox/AgeingExperiment/Paper/Figures/ProductionSupplemental/SF13/Valid_TCRs.pdf",
       height=3.55, width=6.15, useDingbats=FALSE)
```

There's a bit of fluctuation, but for the most part ~70% of simulated TCRs are non-functional. I would expect that in reality when we do the TCR-sequencing, we will have a 
population of thymocytes strongly enriched for functional TCRs (part of cortical positive selection).

I'll subset to just the functional TCRs for each sample size.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
valid.tcr <- tcr.sims[tcr.sims$Valid.TCR %in% c("True"), ]
table(valid.tcr$Sample.Size)
```

Next, what does the distribution of TRA and TRB V(D)J regions look like for each sample size?

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5.25, fig.width=14.75}
trav.table <- as.data.frame(apply(table(valid.tcr$TRA_V, valid.tcr$Sample.Size), 2, FUN=function(X) X/sum(X)))
trav.table$TRA_V <- rownames(trav.table)
trav.table <- melt(trav.table, id.vars='TRA_V')

ggplot(trav.table, aes(x=TRA_V, y=value, fill=variable)) +
  geom_bar(position='dodge', stat='identity') +
  theme_mike() +
  scale_fill_Publication() +
  theme(axis.text.x = element_blank()) +
  labs(x="TRA V segment", y="Proportion") +
  facet_wrap(~variable, scales='free') +
  NULL

ggsave(filename="~/Dropbox/AgeingExperiment/Paper/Figures/ProductionSupplemental/SF13/TRA_V_usage.pdf",
       height=5.25, width=14.75, useDingbats=FALSE)
```

The x-axis are the unorderd TRA V segments. As expected as the sample size increases we start to get beter representtion of V segments. By around 10-20k initial TCRs we have 
a pretty good, and uniform, representation of all segments.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5.25, fig.width=14.75}
traj.table <- as.data.frame(apply(table(valid.tcr$TRA_J, valid.tcr$Sample.Size), 2, FUN=function(X) X/sum(X)))
traj.table$TRA_J <- rownames(traj.table)
traj.table <- melt(traj.table, id.vars='TRA_J')

ggplot(traj.table, aes(x=TRA_J, y=value, fill=variable)) +
  geom_bar(position='dodge', stat='identity') +
  theme_mike() +
  scale_fill_Publication() +
  theme(axis.text.x = element_blank()) +
  labs(x="TRA J segment", y="Proportion") +
  facet_wrap(~variable, scales='free') +
  NULL

ggsave(filename="~/Dropbox/AgeingExperiment/Paper/Figures/ProductionSupplemental/SF13/TRA_J_usage.pdf",
       height=5.25, width=14.75, useDingbats=FALSE)
```

We have a similar picture for the J-segments, it looks like by 10,000 sampled TCRs we've got a good sampling of TRA segments.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5.25, fig.width=14.75}
trbv.table <- as.data.frame(apply(table(valid.tcr$TRB_V, valid.tcr$Sample.Size), 2, FUN=function(X) X/sum(X)))
trbv.table$TRB_V <- rownames(trbv.table)
trbv.table <- melt(trbv.table, id.vars='TRB_V')

ggplot(trbv.table, aes(x=TRB_V, y=value, fill=variable)) +
  geom_bar(position='dodge', stat='identity') +
  theme_mike() +
  scale_fill_Publication() +
  theme(axis.text.x = element_blank()) +
  labs(x="TRB V segment", y="Proportion") +
  facet_wrap(~variable, scales='free') +
  NULL

ggsave(filename="~/Dropbox/AgeingExperiment/Paper/Figures/ProductionSupplemental/SF13/TRB_V_usage.pdf",
       height=5.25, width=14.75, useDingbats=FALSE)
```

The x-axis are the unorderd TRB V segments - there are fewer TRB segments than TRA. Again, sampling is less variable around 10-20k TCRs.

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5.25, fig.width=14.75}
trbj.table <- as.data.frame(apply(table(valid.tcr$TRB_J, valid.tcr$Sample.Size), 2, FUN=function(X) X/sum(X)))
trbj.table$TRB_J <- rownames(trbj.table)
trbj.table <- melt(trbj.table, id.vars='TRB_J')

ggplot(trbj.table, aes(x=TRB_J, y=value, fill=variable)) +
  geom_bar(position='dodge', stat='identity') +
  theme_mike() +
  scale_fill_Publication() +
  theme(axis.text.x = element_blank()) +
  labs(x="TRB J segment", y="Proportion") +
  facet_wrap(~variable, scales='free') +
  NULL

ggsave(filename="~/Dropbox/AgeingExperiment/Paper/Figures/ProductionSupplemental/SF13/TRB_J_usage.pdf",
       height=5.25, width=14.75, useDingbats=FALSE)
```

The same picture for TRB J segments. The TRB D segments were not recorded in the simulations. From these analyses, it looks like somewhere between 10-20k cells should be 
sufficient to accurately estimate the different TCR segment usage statistics.

Now for CDR3 entropy, which is what we are interested in from a repertoire diversity perspective.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(valid.tcr, aes(x=Sample.Size, y=TRA_CDR3_entropy)) +
  geom_boxplot() +
  theme_mike() +
  NULL

ggsave(filename="~/Dropbox/AgeingExperiment/Paper/Figures/ProductionSupplemental/SF13/TRA_CDR3_entropy.pdf",
       height=3.95, width=6.75, useDingbats=FALSE)
```

This is the distribution of TRA CDR3 entropies - it looks pretty stable from ~5000 TCRs onwards.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(valid.tcr, aes(x=Sample.Size, y=TRB_CDR3_entropy)) +
  geom_boxplot() +
  theme_mike() +
  NULL

ggsave(filename="~/Dropbox/AgeingExperiment/Paper/Figures/ProductionSupplemental/SF13/TRB_CDR3_entropy.pdf",
       height=3.95, width=6.75, useDingbats=FALSE)
```

This is the distribution of TRB CDR3 entropies. It also looks stable from ~5000 samples TCRs onwards. What about the whole TCR repertoire entropy? This is a reflection of 
how diversity the whole _pool_ of functional TCRs is. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
aa.alphabet <- c("F", "I", "S", "T", "Y", "N", "C", "L", "K", "\\*", "R", "M", "W", "V", "P", "A", "H", "D", "G", "Q", "E")
calc_freqs <- function(aa, alphabet){
  str_count(aa, alphabet)/sum(str_count(aa, alphabet))
}

calc_entropy <- function(freq.vector){
  require(lettercase)
  # given a vector of frequencies, calculate the Shannon Entropy for each
  # -sum(pi*log(pi)), where pi =  frequency
  # hmm, need to deal with the log(0)s, set Inf to 0
  log.freq <- log2(freq.vector)
  log.freq[is.infinite(log.freq)] <- 0
  
  -sum(freq.vector * log.freq)
}
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# tabulate the CDR3 clonotypes in each sample
tra.clono.tab <- as.data.frame(apply(table(valid.tcr$TRA_CDR3, valid.tcr$Sample.Size), 2, FUN=function(X) X/sum(X)))
tra.cdr3.entropy <- unlist(lapply(tra.clono.tab, FUN=function(A) calc_entropy(A)))

trb.clono.tab <- as.data.frame(apply(table(valid.tcr$TRB_CDR3, valid.tcr$Sample.Size), 2, FUN=function(X) X/sum(X)))
trb.cdr3.entropy <- unlist(lapply(trb.clono.tab, FUN=function(A) calc_entropy(A)))

entropy.df <- data.frame("Sample.Size"=levels(valid.tcr$Sample.Size), "TRA.Entropy"=tra.cdr3.entropy, "TRB.Entropy"=trb.cdr3.entropy)
entropy.melt <- melt(entropy.df, id.vars='Sample.Size')
entropy.melt$Chain <- "Alpha"
entropy.melt$Chain[grepl(entropy.melt$variable, pattern="TRB")] <- "Beta"

entropy.melt$Sample.Size <- factor(entropy.melt$Sample.Size,
                                   levels=c(10, 100, 500, 1000, 5000, 10000, 20000, 50000, 100000))
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(entropy.melt, aes(x=Sample.Size, y=value, fill=Chain)) +
  geom_point(shape=21, size=3.5) +
  theme_mike() +
  scale_fill_Publication() +
  expand_limits(y=c(0)) +
labs(x="Sample Size", y="Shannon Entropy")
```

This is the entropy, using the frequency of distinct clonotypes to measure diversity. I'd also like to know what the distribution or spread of these entropy values are. This 
is where the additional simulations come in which act as replicates; I'll show it with 1-5 replicates.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
col.names <- c("TRA_V",	"TRA_J",	"TRA_len", "TRA_nt", "TRA_entropy", "TRA.Misc", "TRA_AA",	"TRA_Peptide",	"TRA_Valid",	"TRA_CDR1_entropy",	"TRA_CDR1",
               "TRA_CDR2_entropy",	"TRA_CDR2",
               "TRA_CDR3_entropy",	"TRA_CDR3",	"TRB_V",	"TRB_J",	"TRB_len", "TRB_nt",	"TRB_entropy",	"TRB_AA",	"TRB_Peptide",	"TRB_Valid",	"TRB_CDR1_entropy",	"TRB_CDR1",
               "TRB_CDR2_entropy",	"TRB_CDR2",	"TRB_CDR3_entropy",	"TRB_CDR3",	"Valid.TCR")
tcr.dir <- "~/CI_filesystem/mnt/scratchb/jmlab/morgan02/TCR_simulations/results/"
sample.size.files <- list.files(tcr.dir, pattern="TCRsim")
sd.list <- list()

for(x in seq_along(sample.size.files)){
  s.size <- unlist(strsplit(sample.size.files[x], split="-", fixed=TRUE))[1]
  run.number <- gsub(unlist(strsplit(sample.size.files[x], split="-", fixed=TRUE))[3], pattern="\\.txt", replacement="")
  
  s.df <- read.table(paste0(tcr.dir, sample.size.files[x]), sep="\t", header=FALSE, stringsAsFactors=FALSE)
  colnames(s.df) <- col.names
  s.df$Sample.Size <- as.numeric(gsub(s.size, pattern="Sample_", replacement=""))
  s.df$Run <- run.number
  sd.list[[paste0(run.number, s.size)]] <- s.df
}

tcr.sims <- do.call(rbind.data.frame,
                    sd.list)
tcr.sims$Sample.Size <- factor(tcr.sims$Sample.Size,
                               levels=c(10, 100, 500, 1000, 5000, 10000, 20000, 50000, 100000))
valid.reps <- tcr.sims[tcr.sims$Valid.TCR %in% c("True"), ]
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# tabulate the CDR3 clonotypes in each sample
tra.xtab <- xtabs(data=valid.reps, ~ TRA_CDR3 + Sample.Size + Run)
tra.clono.tab <- as.data.frame(tra.xtab)

# normalise within sample size and run number
norm.freq <- by(tra.clono.tab$Freq, INDICES=list("Sample.Size"=tra.clono.tab$Sample.Size, "Run"=tra.clono.tab$Run),
                FUN=function(X) X/sum(X))
tra.clono.tab$NormFreq <- unlist(norm.freq)
tra.rep.entropy <- by(tra.clono.tab$NormFreq, INDICES=list("Sample.Size"=tra.clono.tab$Sample.Size, "Run"=tra.clono.tab$Run),
                      FUN=function(E) calc_entropy(E))

# tabulate the TRB CDR3 clonotypes in each sample
trb.xtab <- xtabs(data=valid.reps, ~ TRB_CDR3 + Sample.Size + Run)
trb.clono.tab <- as.data.frame(trb.xtab)

# normalise within sample size and run number
norm.freq <- by(trb.clono.tab$Freq, INDICES=list("Sample.Size"=trb.clono.tab$Sample.Size, "Run"=trb.clono.tab$Run),
                FUN=function(X) X/sum(X))
trb.clono.tab$NormFreq <- unlist(norm.freq)
trb.rep.entropy <- by(trb.clono.tab$NormFreq, INDICES=list("Sample.Size"=trb.clono.tab$Sample.Size, "Run"=trb.clono.tab$Run),
                      FUN=function(E) calc_entropy(E))

comb.df <- expand.grid(levels(tra.clono.tab$Sample.Size), levels(tra.clono.tab$Run))
combos <- paste(comb.df$Var1, comb.df$Var2, sep="_")

rep.entropy.df <- data.frame("Combo"=combos, "TRA.Entropy"=as.numeric(tra.rep.entropy), "TRB.Entropy"=as.numeric(trb.rep.entropy))
rep.entropy.df$Sample.Size <- as.numeric(unlist(lapply(strsplit(as.character(rep.entropy.df$Combo), split="_", fixed=TRUE), FUN=function(P) paste0(P[1]))))
rep.entropy.df$Run <- unlist(lapply(strsplit(as.character(rep.entropy.df$Combo), split="_", fixed=TRUE), FUN=function(P) paste0(P[2])))
rep.entropy.melt <- melt(rep.entropy.df, id.vars=c('Sample.Size', 'Run', 'Combo'))
rep.entropy.melt$Chain <- "Alpha"
rep.entropy.melt$Chain[grepl(rep.entropy.melt$variable, pattern="TRB")] <- "Beta"
 
rep.entropy.melt$Sample.Size <- factor(rep.entropy.melt$Sample.Size,
                                       levels=c(10, 100, 500, 1000, 5000, 10000, 20000, 50000, 100000))
#rep.entropy.melt$value <- as.numeric(as.character(rep.entropy.melt$value))
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(rep.entropy.melt, aes(x=Sample.Size, y=value, fill=Chain)) +
  geom_boxplot() +
  #geom_point(shape=21, size=3.5) +
  theme_mike() +
  scale_fill_Publication() +
  expand_limits(y=c(0)) +
labs(x="Sample Size", y="Shannon Entropy")

ggsave(filename="~/Dropbox/AgeingExperiment/Paper/Figures/ProductionSupplemental/SF13/TCR_entropy_10reps.pdf",
       height=3.95, width=6.75, useDingbats=FALSE)
```

This shows the distribution of entropyies across all 10 replicates at different sample sizes, what about using different numbers of replicates?

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(rep.entropy.melt[rep.entropy.melt$Run %in% paste0("Run", 1:3), ], aes(x=Sample.Size, y=value, fill=Chain)) +
  geom_boxplot() +
  #geom_point(shape=21, size=3.5) +
  theme_mike() +
  scale_fill_Publication() +
  expand_limits(y=c(0)) +
labs(x="Sample Size", y="Shannon Entropy")

ggsave(filename="~/Dropbox/AgeingExperiment/Paper/Figures/ProductionSupplemental/SF13/TCR_entropy_3reps.pdf",
       height=3.95, width=6.75, useDingbats=FALSE)
```

This is the distribution of Shannon entropies using just 3 replicates. These look pretty consistent. These calculations also give us an idea of the magnitude of the Shannon 
entropy values to expect. So for 10-20k input cells we might expect Shannon entropy values between ~11-13.




